---
title: 'Lab 10: Halloween Candy'
author: 'Nicholas Do (PID: A15053002)'
date: "10/28/2021"
output: html_document
---

```{r}
candy_file <- "candy-data.csv"

candy = read.csv(candy_file, row.names=1)
#head(candy)
candy

```
> Q1. How many different candy types are in this dataset?

```{r}
nrow(candy)
```
There are 85 candies in the dataset. 

> Q2. How many fruity candy types are in the dataset?

```{r}
sum(candy$fruity)
```
There are 38 fruity candy types.

# chocolatey?
```{r}
sum(candy$chocolate)
```

Win percent example
```{r}
candy["Twix", ]$winpercent
```
> Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

```{r}
#Favorite candy: Reeses
candy["ReeseÃ•s Peanut Butter cup", ]$winpercent
```

> Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat", ]$winpercent
```

> Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```
Skimr
```{r}
library("skimr")
skim(candy)
```



> Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

Yes, winpercent is scaled differently. 

> Q7. What do you think a zero and one represent for the candy$chocolate column?

Boolean T/F value for whether or not the candy is chocolate.


> Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent)
```

> Q9. Is the distribution of winpercent values symmetrical?

The distribution of values does not look symmetrical, the center values
have higher frequency than the outer ones.

> 10. Is the center of the distribution above or below 50%?

The center of the distribution is below 50%; it lies at around 20%

> Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
chocowin <- candy$winpercent[as.logical(candy$choco)]
mean(chocowin)

fruitywin <- candy$winpercent[as.logical(candy$fruity)]
mean(fruitywin)

```
On average, chocolate candy is ranked higher than fruit candy


> Q12. Is this difference statistically significant?

```{r}
t.test(chocowin, fruitywin)
```

Yes, it is statistically significant!



# 3. Overall Candy Rankings

> Q13. What are the five least liked candy types in this set?


```{r}
head(candy[order(candy$winpercent),], n=5)
```


> Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
head(candy[order(candy$winpercent, decreasing=TRUE),], n=5)
```



> Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
library(ggplot2)

ggplot(candy) + 
  aes(winpercent, rownames(candy)) +
  geom_col()
```

> Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy), winpercent)) +
  geom_col()
```



Add colors
```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"

ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```


> Q17. What is the worst ranked chocolate candy?

Worst ranked chocolate candy is sixlets.

> Q18. What is the best ranked fruity candy?

The best ranked fruity candy is starburst.


#4. Looking at pricepercent

```{r}
library(ggrepel)

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```


> Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

Reese's Miniatures seems to be the best bang for our buck, it has the highest winpercent among
the low cost candies.


> Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?


```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```


> Q21. Make a barplot again with geom_col() this time using pricepercent and then improve this step by step, first ordering the x-axis by value and finally making a so called “dot chat” or “lollipop” chart by swapping geom_col() for geom_point() + geom_segment().



```{r}
ggplot(candy) + 
  aes(pricepercent, rownames(candy)) +
  geom_col()

ggplot(candy) + 
  aes(pricepercent, reorder(rownames(candy), pricepercent)) +
  geom_col()

# Make a lollipop chart of pricepercent
ggplot(candy) +
  aes(pricepercent, reorder(rownames(candy), pricepercent)) +
  geom_segment(aes(yend = reorder(rownames(candy), pricepercent), 
                   xend = 0), col="gray40") +
    geom_point()
```

# 5. Exploring correlation structure

```{r}
library(corrplot)

cij <- cor(candy)
corrplot(cij)
```

> Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

Fruity and chocolatey have the biggest anti-correlation. Pluribus and bar also have a large
anti-correlation. In general, the orange colored relationships have an anti-correlation
as observed from the scale on the right of the matrix.

> Q23. Similarly, what two variables are most positively correlated?

The two variables that are most positively correlated are chocolate with winpercent and bar.

# 6. PCA

```{r}
#pca_test <- prcomp(candy)
pca <- prcomp(candy, scale = TRUE)
#summary(pca_test)
summary(pca)

```

```{r}
plot(pca$x[,1:2])
```

Add color
```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```

Make a better plot with ggplot2

```{r}
# Make a new data-frame with our PCA results and candy data
my_data <- cbind(candy, pca$x[,1:3])



p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

p

#Use ggrepel to add labels
p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")

```



```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

> Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

fruity, hard, and pluribus 
