---
title: 'Lab 17: Vaccination Rates'
author: 'Nicholas Do (PID: 15053002)'
date: "11/28/2021"
output: pdf_document
---

Read the CSV

```{r}
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
head(vax)

```

>> Q1. What column details the total number of people fully vaccinated?

```{r}
head(vax[,9])
head(vax$persons_fully_vaccinated)
```


Column 9, or vax$persons_fully_vaccinated represents the 
column for persons fully vaccinated.


>> Q2. What column details the Zip code tabulation area? 

```{r}
head(vax$zip_code_tabulation_area)
```
 
>> Q3. What is the earliest date in this dataset?

```{r}
min(vax$as_of_date)
max(vax$as_of_date)
```
The earliest date is 01/05/2021

Q4. What is the latest date in this dataset?

The latest date is 11/23/2021


```{r}
#check for skimr package
library(skimr)
```

```{r}
#skimr::skim(vax)
```
>> Q5. How many numeric columns are in this dataset?


There are 9 numeric columns in the dataset.

>> Q6. Note that there are “missing values” in the dataset. 
How many NA values there in the persons_fully_vaccinated column?

There are 8355 missing values in the persons_fully_vaccinated column


>> Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)?

```{r}
p <- 8355 / 82908 
p
```

10.08% missing


>> Q8. [Optional]: Why might this data be missing?

This data is probably missing because those individuals are not fully vaccinated yet.


#Working with dates

```{r}
library(lubridate)
```

```{r}
#Today's date
today()
```



```{r}
#Convert to lubridate
#Specify that we are using the Year-mont-day format
vax$as_of_date <- ymd(vax$as_of_date)
```



```{r}
#How many days have passed since the first vaccination reported?
today() - vax$as_of_date[1]
```


```{r}
#How many days does the dataset span?
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```


>> Q9. How many days have passed since the last update of the dataset?

```{r}
today() - max(vax$as_of_date)
```
It has been 5 days since the last update of the dataset.


>> Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)?


```{r}
unique(vax$as_of_date)
```

There are 47 unique dates in the dataset.



#Working with ZIP codes


```{r}
library(zipcodeR)
```
```{r}
geocode_zip('92037')

#Distance between zipcodes

zip_distance('92037','92109')
```

```{r}
reverse_zipcode(c('92037', "92109") )
```


#Focus on SD Area

```{r}
library(dplyr)

sd <- filter(vax, county == "San Diego")

nrow(sd)
```

```{r}
sd.10 <- filter(vax, county == "San Diego" &
                age5_plus_population > 10000)
```


>> Q11. How many distinct zip codes are listed for San Diego County?


```{r}
length(unique(sd$zip_code_tabulation_area))
```

There are 107 unique zip codes in San Diego County


>> Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset?


```{r}
which.max(sd$age12_plus_population)

sd$zip_code_tabulation_area[60]

```
92154 is the zip code that has the largest 12+ population in this dataset.


```{r}
anothersd <- filter(vax, county == "San Diego", as_of_date == "2021-11-09")
nrow(anothersd)
```


```{r}
head(anothersd[,1])
```
```{r}
#skimr::skim(anothersd)
```


>> Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2021-11-09”?


```{r}
anothersd$percent_of_population_fully_vaccinated
```
0.963 or 96.3%



>> Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2021-11-09”

```{r}
hist(anothersd$percent_of_population_fully_vaccinated)
```






#Focus on UCSD/La Jolla

```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```

>> Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:


```{r}
library(ggplot2)
ggplot(ucsd) +
  aes(ucsd$as_of_date,
      ucsd$percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x = "Date", y="Percent Vaccinated")
```


#Compare to other similar areas

```{r}
# Subset to all CA areas with a population as large as 92037
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2021-11-16")

head(vax.36)
```

>> Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?

```{r}
vaxmean <- mean(vax.36$percent_of_population_fully_vaccinated)

p <- ggplot(ucsd) +
  aes(ucsd$as_of_date,
      ucsd$percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_hline(yintercept = vaxmean, linetype = "dashed") +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x = "Date", y="Percent Vaccinated")
p

```

>> Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”?

```{r}
summary(vax.36)
```



>> Q18. Using ggplot generate a histogram of this data.


```{r}
ggplot(vax.36) + aes(vax.36$percent_of_population_fully_vaccinated) + geom_histogram()
```


>> Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?


```{r}
vax %>% filter(as_of_date == "2021-11-16") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)

vax %>% filter(as_of_date == "2021-11-16") %>%  
  filter(zip_code_tabulation_area=="92109") %>%
  select(percent_of_population_fully_vaccinated)
```

92040 is under, 92109 is above.


>> Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144.


```{r}
vax.36.all <- filter(vax, age5_plus_population > 36144)


ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="blue") +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated",
       title="Vaccination rate across California",
       subtitle="Only areas with a population above 36k are shown.") +
  geom_hline(yintercept = vaxmean, linetype="dashed")
```


>> Q21. How do you feel about traveling for Thanksgiving and meeting for in-person class next Week?

I feel comfortable with it due to the high vaccination rate across California, according to this dataset. 







