---
title: 'Lab 11: Structural Bioinformatics II'
author: 'Nicholas Do (PID: 15053002)'
date: "11/14/2021"
output: pdf_document
---


# 1.1 Obtaining and inspecting our input structure
```{r}
library(bio3d)


file.name <- get.pdb("1hsg")
hiv <- read.pdb(file.name)
```



```{r}
hiv
```
> Q1: What is the name of the two non protein resid values in this structure? What does resid
correspond to and how would you get a listing of all reside values in this structure? 

The two non protein resid values are HOH and MK1. Resid corresponds
to the residues within the structure. To get a list of all residue
values in the structure look at the variable through: hiv$atom$resid


```{r}
head(hiv$atom$resid, n = 20)
```


# 1.2 Prepare initial protein and ligand input files



```{r}
prot <- trim.pdb(hiv, "protein")
lig <- trim.pdb(hiv, "ligand")

write.pdb(prot, file="1hsg_protein.pdb")
write.pdb(lig, file="1hsg_ligand.pdb")
```


# 1.3 Using AutoDockTools to setup protein docking input


> Q2: Can you locate the binding site visually? Note that
crystal structures normally lack hydrogen atoms, why? 


Yes, you can locate the binding site visually. The hydrogen atoms
are not generally required to represent the structure, and 
normally crystal structures lack hydrogen atoms. 


> Q3: Look at the charges. Does it make sense (e.g. based on your knowledge of the
physiochemical properties of amino acids)?

Yes,the charges make sense




















