---
title: "Lab 6: R Functions"
author: "Nicholas Do (PID: A15053002)"
date: "10/14/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R functions

In today's class we are going to write a function together that grades some students' work.

> **Q1**.Write a function grade() to determine an overall grade from a vector of student homework assignment scores dropping the lowest single score. If a student misses a homework (i.e. has an NA value) this can be used as a score to be potentially dropped. Your final function should be adquately explained with code comments and be able to work on an example class gradebook such as this one in CSV format: “https://tinyurl.com/gradeinput” [3pts]


Example input vectors to start with:
```{r}
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)
```

Let's start with student1 and find their average score.
```{r}
mean(student1)
```

But we want to drop the lowest score...
We could try the **min()** function

```{r}
min(student1)
```



The **which.min()** function looks more useful:
```{r}
minVal <- which.min(student1)
student1[minVal]
```
Then, we calculate the mean with the remaining values.
```{r}
gradeVector <- student1[-minVal]
gradeVector
average <- mean(gradeVector)
average
```

#For student2
```{r}
student2
```

However, it gives NA if a student has a missing homework
```{r}
mean(student2[-which.min(student2)])
```

We need to remove the NA elements of the vector

```{r}
mean(student2, na.rm=TRUE)
#mean(student2[ -which.min(student2) ], na.rm = TRUE)
#Not what we want, drops 80 and not the NA
```

Let's look at student3
```{r}
student3
```
One new idea/approach is we could replace the NA (missing homeworks) with zero. 

Let's try with student2

```{r}
student2

naIndex <- which(is.na(student2))
naIndex

student2[naIndex] <- 0
student2

student2 <- student2[-which.min(student2)]
student2

```



Lets try with student3:
```{r}
student3
temp <- student3
naVector<- which(is.na(student3))
naVector
temp[naVector] <- 0
temp
temp <- temp[-which.min(temp)]
temp
mean(temp)
```

```{r}
grades <- function(x) {
  #Make sure our scores are all numbers
  currentGrades <- as.numeric(x)
  currentGrades[ which(is.na(currentGrades)) ] <- 0
  average <- mean(currentGrades[ -which.min(currentGrades) ])
  average
  

}
```

```{r}
testVector <- c(100, 100, 100, 60, 40, 20)
grades(testVector)
#student1
grades(student1)
#student2
grades(student2)
#student3
grades(student3)
```





Now read the full gradebook CSV file.
```{r}
scores <- read.csv("https://tinyurl.com/gradeinput", row.names=1)
scores
```


```{r}
#grades(scores[1,])
finalScores <- apply(scores, 1, grades)

```


> **Q2** Using your grade() function and the supplied gradebook, Who is the top scoring student
overall in the gradebook? [3pts]

The top scoring student is student 18.
```{r}
which.max(finalScores)
```

> **Q3** From your analysis of the gradebook, which homework was toughest on students (i.e. obtained the lowest scores overall? [2pts]

Apply to the columns
```{r}
apply(scores, 2, mean, na.rm=TRUE)
```
HW 3 was the worst



